---
title: "Time Series Project"
author: "Daniel Fields & Max Gershman & Luis Aragon"
date: "2/13/2019"
output: html_document
---

**Monthly Retail Sales: Beer, Wine, and Liquor Stores** 

- Units:  Millions of Dollars, Not Seasonally Adjusted
- Frequency:  Monthly

_*Suggested Citation:*_

*U.S. Bureau of the Census, Retail Sales: Food and Beverage Stores [MRTSSM445USN], retrieved from FRED, Federal Reserve Bank of St. Louis; https://fred.stlouisfed.org/series/MRTSSM445USN, February 16, 2019.*

```{r}
#For
install.packages("tseries",repos = "https://cloud.r-project.org/") 
#For
install.packages("tidyverse",repos = "https://cloud.r-project.org/")
#For
install.packages("forecast",repos = "https://cloud.r-project.org/")
#For
install.packages("plotrix", repos = "https://cloud.r-project.org/")
#For
install.packages("astsa",repos = "https://cloud.r-project.org/" )
#For 
install.packages("MASS",repos = "https://cloud.r-project.org/" )
#For
install.packages("TSA",repos = "https://cloud.r-project.org/" )
#For
install.packages("GeneCycle",repos = "https://cloud.r-project.org/" )
#For
install.packages("zoo",repos = "https://cloud.r-project.org/")
install.packages("xts",repos = "https://cloud.r-project.org/")
#For Leven Test on Homoskedacity
install.packages("car")
```

```{r}
library(tseries)
library(tidyverse)
library(forecast)
library(plotrix)
library(MASS)
library(astsa)
library(GeneCycle)
library(TSA)
library(car)
```
### Load Data and turn into Time Series Object
```{r}
booze <- read.csv("booze.csv", header = TRUE)
str(booze)
head(booze)

#Convert Data to time series
booze <- ts(booze[,2],frequency = 12, start = c(1992,1)) #Monthly data points
var_og <- var(booze)
```

#Stabilizing Variance
```{r}
#Plot raw time series
ts.plot(booze, main = "Monthly Retail Sales: Beer, Wine, and Liquor Stores" )

time <- 1:length(booze)
lin.model <- lm(booze ~ time)

predicted <- predict(lin.model)
residuals <- residuals(lin.model)
xy <- data.frame(time,booze,predicted,residuals)

ggplot(xy, aes(x = time, y = booze)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
  geom_segment(aes(x = time,y = booze,xend = time, yend = predicted), alpha = .5) +
  geom_point(aes(x = time, y = booze, fill = "Actual"), size = 0.4)+
  ggtitle("Linear Model Intuition for Slight Heteroskedacity")

qqnorm(residuals, pch = 1, frame = FALSE)
qqline(residuals, col = "steelblue", lwd = 2)
```
We suspect this time series may be heteroskedastic so to test this employ a Levene Test. Per R Documentation, the Levene Test is a Bayesian form of Levene's test. Levene's test is used to assess the probability of the equality of residual variances in different groups. 
So then, we observe a small degree of heteroskedacity. To fix this, we will apply a Box-Cox transform to the data. A Box-Cox transform, where we apply a transform to the series based on $\lambda$ to the original data $U_t$.
$$f_{\lambda}(U_t) = \begin{array}{cc}
  \{ & 
    \begin{array}{cc}
      ln(U_t) & \lambda = 0 \\
      \lambda^{-1}(U_t^{\lambda} - 1) & \lambda  \neq 0\\
    \end{array}
\end{array} $$
```{r}
#Do Box-Cox
transforms <- boxcox(booze~c(1:length(booze)))
which_transform <- data.frame(transforms)%>%
  filter(y == max(y))

which_transform <- which_transform$x
```

After doing this, we get that a Box-Cox transform according to the best value $\lambda_{B}=$ `r which_transform` will work best. Now we will apply the box cox to the transform to the data. Since our best value for $\lambda_B$ was slightly less than 1, then our transformation will be $\frac{(U_t^{\lambda_B} - 1)}{\lambda_B}$, where $U_t$ is the booze time series data.
```{r}
BoxCox <- function(data, lambda){
  data_transformed <- (data^lambda - 1)/lambda
  return(data_transformed)
}
#Transform the time series
booze2 <- BoxCox(booze, which_transform)

#Examine the effects
ts.plot(booze2, main = "Monthly Retail Sales: Beer, Wine, and Liquor Stores After Box-Cox Transform" )

time <- 1:length(booze2)
lin.model2 <- lm(booze2 ~ time)

predicted2 <- predict(lin.model2)
residuals2 <- residuals(lin.model2)
xy2 <- data.frame(time,booze2,predicted2,residuals2)

ggplot(xy2, aes(x = time, y = booze2)) +
  geom_smooth(method = "lm", se = FALSE, color = "lightgrey") +
  geom_segment(aes(x = time,y = booze2,xend = time, yend = predicted2), alpha = .5) +
  geom_point(aes(x = time, y = booze2, fill = "Actual"), size = 0.4)+
  ggtitle("Linear Model After Box-Cox")+
  ylab("Price Transformed")

var_boxcox <- var(booze2)


```

```{r}
decomposed_model <- decompose(booze2, type =  "additive")

autoplot(decomposed_model, main = "Additive Decomposition Plot")+
  xlab("Time in Months")
#Examine the data for seasonality
seasonplot(booze2, 12, col = rainbow(12),year.labels = TRUE, main = "Annual Seasonality Plot")
```
### Remove Seasonality and Trend

We observe a seasonality of s = 12, and a linear trend (of order 1).
So, let $X_t$ be the Box-Cox transformed time series of retail sales of alchohol in millions of dollars.
So then, $\nabla \nabla_{12} X_t \text{ Should be a stationary time series}$.

```{r removeTrend_Seasonality}
head(booze2)
var(booze2)
ts.plot(booze2, main = "Box-Cox Transformed Data")

#Remove Seasonality
booze_diff_seasonality <- diff(booze2, lag = 12)
ts.plot(booze_diff_seasonality, main = expression(paste(nabla[12],X[t])), ylab = "Seasonality Removed")
abline(lm(booze_diff_seasonality ~ as.numeric(1:length(booze_diff_seasonality))))
var_season_removed <- var(booze_diff_seasonality)


#Remove trend
booze_stationary <- diff(booze_diff_seasonality, lag = 1)
ts.plot(booze_diff_seasonality, main = expression(paste(nabla,nabla[12],X[t])))
var_stationary_diff1 <- var(booze_stationary)

#Check if more differencing helps
once_more <- diff(booze_stationary, lag = 1)
var_stationary_diff2 <- var(once_more)
two_more <- diff(once_more, lag = 1)
var_stationary_diff3 <- var(two_more)
variances <- c(var_boxcox,var_season_removed,var_stationary_diff1,var_stationary_diff2, var_stationary_diff3)
labels <- c( "After Box Cox", "After Removing Seasonality", "After Removing Trend", "After Differencing Stationary Once", "After Differencing Stationary Twice")

data.frame(variances, labels)%>%
  ggplot(aes(x = labels, y = variances, fill = variances))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

So, after removing seasonality, the variance is lowest. So then, this series is `r booze_diff_seasonality`

### Model Building

```{r}
par(cex.main = 0.8)
booze_data <- ts(booze_diff_seasonality,frequency = 12, start = c(1992,1)) #Monthly data points
#Split data into train and test sets.
booze_train <- head(booze_diff_seasonality, length(booze_diff_seasonality) - 30)
booze_test <- tail(booze_diff_seasonality,30)
Acf(booze_train, lag.max = 60, main = "Sample ACF")
Pacf(booze_train,lag.max = 60, main = "Sample PACF")#MAKE NOTE WHY CHANGE AXIS
```

We will observe annual seasonality.
```{r}
decomposed_model_train <- decompose(booze_train, type =  "additive")

autoplot(decomposed_model_train, main = "Additive Decomposition Plot")+
  xlab("Time in Months")
```
